---
title: "vignette-ggradial"
author: "Thorben Hebbelmann, Marcel Schulte, Michael Thane"
date: "03/01/2022"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette-ggradial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
body {
text-align: justify}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

The package 'ggradial' provides wrapper functions for ‘ggplot2’ with radial bar chart and radial line chart extensions in order to visualize high-dimensional clusters and their evolution.

* radial_barchart_static: Creates a radial barchart which shows a cluster's average feature values.
* radial_barchart_compare: Creates a radial barchart which shows a cluster's average feature values and allows a comparison between different points in time. 
* radial_line_chart: Creates a radial line chart which shows the average feature values for each cluster.


## Installation

The package is hosted on [GitHub] and can be installed via: 

```{r install-github, echo=TRUE, eval=FALSE}
devtools::install_github("https://github.com/mthane/ggradial",auth_token="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")
```

[GitHub]: https://github.com/mthane/ggradial

```{r libraries, message=FALSE, warning=FALSE}
  library(ggradial)
  library(tidyverse)
  library(ggiraph)
  library(kableExtra)
```
## Usage Examples

First we take a view over the parameters and their functions:

* df: The dataset as data frame which should be scaled and only contain the category values in a wide table form.
* group_names: Character vector with group names of features. The number of columns in df and the length of group_names must be equal. Group names will be displayed in the inner circle.
* inner_label: Label inside the inner circle, e.g. name of the chart (optional).
* color_inner_circle Color of the inner circle (optional).
* scale_rng Min and max values to be shown (optional).
* interactive Boolean to indicate whether the plot should be interactive or not (optional).
* tooltip_bars: String to choose which tooltip texts will be displayed for the bars (when interactive). Options to select from: average, sd (standard deviation), ci (confidence interval), all. Default is average (optional).
* tooltip_labels: Data frame to provide the tooltip texts for the labels (when interactive), needs to have a "label" and a "description" column. The description provides the tooltip text for the assigned label (optional).
* show_group_names: Boolean to indicate whether the group names should be shown inside the inner circle. Default is true (optional).

Only comparison chart:

* id: String containing the column name of the ids in df (optional).
* phase: String containing the column name of the phase in df (optional).
* feature_names: Column names of all the features (optional).
* delta_threshold: Numeric value that must be exceeded to show that the treatment had an effect. Must be between 0.05 and 1.0 (optional).

Only line chart:

* cluster: String containing the column name, which indicates the cluster membership in df (optional).
* color_clusters: Colors of the different clusters (optional).


Note: It is recommended to use more than 30 features for an optimal visualization. The interactive mode for the radial_barchart_compare is only available for 2 treatments.

In the following table we can see the columns .id, .phase and .cluster.
In this data set there are patients with an id and they belong to a certain cluster. On top of that there are two phases of a treatment which can be compared in a visualization. With the library it is also possible to compare multiple of such treatments.

## Data Set

```{r dataset, message=FALSE}

DATASET$data%>%
  head()%>%
  kable() %>%
  kable_styling("striped", full_width = F) %>% 
 scroll_box(width = "700px", height = "200px")
```
## Bar Chart

Below we can see a bar chart that shows the average value of a certain numerical feature for each column. In the inner circle we can see the different groups that those features belong to. The groups are saved in DATASET$feature_groups. With the interactive option an interactive plot is drawn. 
The barchart provides a graphical representation of a subpopulation. Since each feature was scaled (z-score normalized) before clustering, it is thus possible to show features with originally different scales in a combined visualization. Due to feature scaling, bars inclined to the outside represent feature averages above the overall population mean and bars inclined to the inside represent feature averages below the population mean. 
For example, if the bar or point of a feature is +1.0, it means that the average feature value is one standard deviation above the feature average of the whole population and a value of -1 indicates that the subpopulation average is 1 standard deviation smaller than the overall population average. 
Furthermore with the comparison bar chart it is possible to compare a subpopulation over time. The color of the arrows and feature labels allows to illustrate the effect of a measurement.

```{r barchart-static, message=FALSE}

dfs = DATASET$data
grs = DATASET$feature_groups

df = dfs[c(".id", ".phase", ".cluster", grs$feature[grs$.gr_id != 1])]

df <- df %>%
  mutate(.cluster = factor(.cluster)) %>%
  filter(.cluster == levels(.cluster)[2L])

data_cluster <- df %>% mutate(.cluster = fct_drop(as.factor(.cluster))) %>% select(-c(.cluster))

group_names = grs$.gr_name[grs$.gr_id != 1]

data_static = data_cluster %>%select(!c(".id",".phase"))
radial_barchart_static(data_static, group_names,interactive = T)
```
## Line Chart
Finally, the radial line chart depicts cluster averages for each feature and provides therefore a compact summary of one or all clusters.


```{r, message=FALSE}

data = DATASET
dfs = data$data
grs= data$feature_groups

group_names = grs$.gr_name[grs$.gr_id != 1]

df = dfs[c(".id", ".phase", ".cluster", grs$feature[grs$.gr_id != 1])]

df <- df %>%
  mutate(.cluster = factor(.cluster)) %>%
  filter(.cluster == levels(.cluster)[2L])
data_linechart = df%>%select(!c(".phase"))

radial_line_chart(data_linechart,group_names, interactive = T,inner_label="Costum Label")
```
## Bar Chart Comparison

Here we can see an example with more then two treatments.


```{r, message=FALSE}
mock = create_mockdata(Nf=30,n_treatments =4,ngroups = 5)
dfs = mock$dfs
grs = mock$grs

df = dfs[c(".id", ".phase", ".cluster", grs$feature[grs$.gr_id != 1])]

df <- df %>%
  mutate(.cluster = factor(.cluster)) %>%
  filter(.cluster == levels(.cluster)[2L])

data_cluster <- df %>% mutate(.cluster = fct_drop(as.factor(.cluster))) %>% select(-c(.cluster))

group_names = grs$.gr_name[grs$.gr_id != 1]

radial_barchart_compare(data_cluster,group_names, interactive = T,inner_label="Costum Label")
```
## Heart Disease Dataset


```{r heart-disease, message=FALSE, warning=FALSE}

# with the heart data set
x <- readr::read_file('http://archive.ics.uci.edu/ml/machine-learning-databases/heart-disease/hungarian.data')
df <- read.table(text = gsub('(\\d)\\n', '\\1 ', x))

radial_barchart_static(df[,-76], "",interactive = T)
```

